---
title: "CodeBook"
author: "John"
date: "10/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Code Book

### Overview

There are five types of Georgia public school data gathered in this project:

 * district and institution names and identifiers
 * details about certified personnel
 * enrollment numbers both total and by subgroup
 * economic status
 * Georgia Milestone outcomes

All data is downloaded from the Georgia Governor's Office of Student Achievement [Downloadable Data](https://gosa.georgia.gov/report-card-dashboards-data/downloadable-data) page.

### Raw Data

#### Certified personnel
Certified personnel files are retrieved for school calendar years 2017-2018 and 2018-2019.
The retrieved data is formatted as CSV files.

|Variable Name|Variable Label| Data Type |
|-------------|--------------|-----------|
| LONG_SCHOOL_YEAR | Long School Year | Character |
| SCHOOL_DSTRCT_CD | School District Code | Character |
| SCHOOL_DSTRCT_NM | School District Name | Character |
| INSTN_NUMBER | Institution Number | Character |
| INSTN_NAME | Institution Name (e.g., Awtry Middle School) | Character |
| DATA_CATEGORY | Data Category | Character |
| DATA_SUB_CATEGORY | Data Subcategory | Character |
| EMPLOYEE_TYPE | Employee Type | Character |
| MEASURE | Measure | Numeric |

Long school year is of the form YYYY-YY, e.g., 2018-19

School District Code is generally a three digit number. "ALL" is included for
summary statistics.

School district name is generally the county name or city name (e.g., Cobb County, Merietta City).

Institution code is generally a thre or four digit number. "ALL" is included for summary
statistics.

Institution name are short names like "Kell High School".  Some institutions are
non-schools, e.g., "Central Office."

Data category differentiates between the various types of records. Data category
has the following values:

 * Certificate Level
 * Certified Personnel
 * Gender
 * Personnel
 * Positions
 * Race/Ethnicity
 * Years Experience
 
Data subcategory takes on different values depending on the data category.  The 
following table lists these values and what the measure denotes.

| Data Category | Data Subcategory Values | Measure |
|---------------|-------------------------|---------|
| "Certificate Level" | "4 Yr Bachelor's", "5 Yr Master's", "6 Yr Specialist's", "7 Yr Doctoral", "Other *"| Count |
| "Certified Personnel" | Provisional, Professional | Count |
| Gender | Male, Female | Count |
| Personnel | Full-time, Part-time | Count |
| Positions | "Average Annual Salary", "Average Contract Days", "Average Daily Salary", Number | Count for Number, Numeric for the rest |
| Race/Ethnicity | Number, Black, Hispanic, Multiracial, "Native America", White | Count |
| "Years Experience" | "< 1", 1-10, 10-20, 20-30, "> 30", Average | Count for factors, but numeric for average |

Employee type differentiates between three types of employees: Administrators,
"PK-12 Teachers", "Support Personnel"

##### Enrollment
Enrollment files are retrieved for school calendar years 2017-2018 and 2018-2019.
The retrieved data is formatted as CSV files. 

|Variable Name|Variable Label| Data Type |
|-------------|--------------|-----------|
| LONG_SCHOOL_YEAR | Long School Year | Character |
| DETAIL_LVL_DESC | Detail Level Description | Character |
| SCHOOL_DSTRCT_CD | School District Code | Character |
| SCHOOL_DSTRCT_NM | School District Name | Character |
| INSTN_NUMBER | Institution Number | Character |
| INSTN_NAME | Institution Name (e.g., Awtry Middle School) | Character |
| GRADES_SERVED_DESC | Grades Served Description | Character |
| ENROLLMENT_PERIOD | Enrollment Period | Character |
| GRADE_LEVEL | Grade Level | Character |
| ENROLLMENT_COUNT | Enrollment Count | Integer |

I ignore everything but those rows with a detail level description of 
"School".

Long school year is of the form YYYY-YY, e.g., 2018-19.

Grades served description is a string of comma separated grade levels (e.g., "PK,KK,01,02")

Enrollment period is one of: Fall, Spring
Grade level is one of: PK, KK, 01, 02, 03, 04, 05, 06, 07, 08, 09, 10, 11, 12.

##### Enrollment by Subgroup
Enrollment by subgroup files are retrieved for school calendar years 2017-2018 and 2018-2019.
The retrieved data is formatted as CSV files. Note that subgroup names differ from
those used in Georgia Milestones End-of-Grade and End-of-Course assessments.

|Variable Name|Variable Label| Data Type |
|-------------|--------------|-----------|
| DETAIL_LVL_DESC | Detail Level Description | Character |
| INSTN_NUMBER | Institution Number | Character |
| SCHOOL_DSTRCT_CD | School District Code | Character |
| LONG_SCHOOL_YEAR | Long School Year | Character |
| INSTN_NAME | Institution Name (e.g., Awtry Middle School) | Character |
| SCHOOL_DSTRCT_NM | School District Name | Character |
| GRADES_SERVED_DESC | Grades Served Description | Character |
| ENROLL_PERCENT_ASIAN | Enrollment Percent ASIAN | | |
| ENROLL_PERCENT_NATIVE | Enrollment Percent NATIVE | Numeric  |
| ENROLL_PERCENT_BLACK | Enrollment Percent BLACK | Numeric  |
| ENROLL_PERCENT_HISPANIC | Enrollment Percent HISPANIC | Numeric  |
| ENROLL_PERCENT_MULTIRACIAL | Enrollment Percent MULTIRACIAL | Numeric  |
| ENROLL_PERCENT_WHITE | Enrollment Percent WHITE | Numeric  |
| ENROLL_PERCENT_MIGRANT | Enrollment Percent MIGRANT | Numeric  |
| ENROLL_PERCENT_ED | Enrollment Percent ED | Numeric  |
| ENROLL_PERCENT_SWD | Enrollment Percent SWD | Numeric  |
| ENROLL_PERCENT_LEP | Enrollment Percent LEP | Numeric  |
| ENROLL_COUNT_REMEDIAL_GR_6_8 | Enrollment Count REMEDIAL_GR_6_8 | Numeric  |
| ENROLL_PCT_REMEDIAL_GR_6_8 | Enrollment Percent REMEDIAL_GR_6_8 | Numeric  |
| ENROLL_COUNT_EIP_K_5 | Enrollment Count EIP_K_5 | Numeric  |
| ENROLL_PERCENT_EIP_K_5 | Enrollment Percent EIP_K_5 | Numeric  |
| ENROLL_COUNT_REMEDIAL_GR_9_12 | Enrollment Count REMEDIAL_GR_9_12 | Numeric  |
| ENROLL_PCT_REMEDIAL_GR_9_12 | Enrollment Percent REMEDIAL_GR_9_12 | Numeric  |
| ENROLL_COUNT_SPECIAL_ED_K12 | Enrollment Count SPECIAL_ED_K12 | Numeric  |
| ENROLL_PCT_SPECIAL_ED_K12 | Enrollment Percent SPECIAL_ED_K12 | Numeric  |
| ENROLL_COUNT_ESOL | Enrollment Count ESOL | Numeric  |
| ENROLL_PCT_ESOL | Enrollment Percent ESOL | Numeric  |
| ENROLL_COUNT_SPECIAL_ED_PK | Enrollment Count SPECIAL_ED_PK | Numeric  |
| ENROLL_PCT_SPECIAL_ED_PK | Enrollment Percent SPECIAL_ED_PK | Numeric  |
| ENROLL_COUNT_VOCATION_9_12 | Enrollment Count VOCATION_9_12 | Numeric  |
| ENROLL_PCT_VOCATION_9_12 | Enrollment Percent VOCATION_9_12 | Numeric  |
| ENROLL_COUNT_ALT_PROGRAMS | Enrollment Count ALT_PROGRAMS | Numeric  |
| ENROLL_PCT_ALT_PROGRAMS | Enrollment Percent ALT_PROGRAMS | Numeric  |
| ENROLL_COUNT_GIFTED | Enrollment Count GIFTED | Numeric  |
| ENROLL_PCT_GIFTED | Enrollment Percent GIFTED | Numeric  |
| ENROLL_PERCENT_MALE | Enrollment Percent MALE | Numeric  |
| ENROLL_PERCENT_FEMALE | Enrollment Percent FEMALE | Numeric  |

I ignore everything but those rows with a detail level description of 
"School".

Long school year is of the form YYYY-YY, e.g., 2018-19.

Grades served description is a string of comma separated grade levels (e.g., "PK,KK,01,02")

##### Georgia Milestones End-of-Grade Assessments
Georgia Milestones End-of-Grade Assessments files are retrieved for school calendar years 2017-2018 and 2018-2019.
End of grade assessments are performed for grades 3 through 8.
The retrieved data is formatted as CSV files. Note that subgroup names differ from
those used in Enrollment by Subgroup.

|Variable Name|Variable Label| Data Type |
|-------------|--------------|-----------|
| LONG_SCHOOL_YEAR | Long School Year | Character |
| SCHOOL_DSTRCT_CD | School District Code | Character |
| SCHOOL_DSTRCT_NM | School District Name | Character |
| INSTN_NUMBER | Institution Number | Character |
| INSTN_NAME | Institution Name (e.g., Awtry Middle School) | Character |
| ACDMC_LVL | Academic Level | Character |
| SUBGROUP_NAME | Subgroup Name | Character |
| TEST_CMPNT_TYP_NM | Test Component Type Name | Character |
| NUM_TESTED_CNT | Number Tested Count | Numeric  |
| BEGIN_CNT | Beginning Count | Numeric  |
| DEVELOPING_CNT | Developing Count | Numeric  |
| PROFICIENT_CNT | Proficient Count | Numeric  |
| DISTINGUISHED_CNT | Distinguished Count | Numeric  |
| BEGIN_PCT | Beginning Percent | Numeric  |
| DEVELOPING_PCT | Developing Percent | Numeric  |
| PROFICIENT_PCT | Proficient Percent | Numeric  |
| DISTINGUISHED_PCT | Distinguished Percent | Numeric  |

##### Georgia Milestones End-of-Course Assessments
Georgia Milestones End-of-Course Assessments files are retrieved for school calendar years 2017-2018 and 2018-2019.
End of course assessments are performed for grades 9 through 12.
The retrieved data is formatted as CSV files. Note that subgroup names differ from
those used in Enrollment by Subgroup.

|Variable Name|Variable Label| Data Type |
|-------------|--------------|-----------|
| LONG_SCHOOL_YEAR | Long School Year | Character |
| SCHOOL_DSTRCT_CD | School District Code | Character |
| SCHOOL_DSTRCT_NM | School District Name | Character |
| INSTN_NUMBER | Institution Number | Character |
| INSTN_NAME | Institution Name (e.g., Awtry Middle School) | Character |
| ACDMC_LVL | Academic Level | Character |
| SUBGROUP_NAME | Subgroup Name | Character |
| TEST_CMPNT_TYP_NM | Test Component Type Name | Character |
| NUM_TESTED_CNT | Number Tested Count | Numeric  |
| BEGIN_CNT | Beginning Count | Numeric  |
| DEVELOPING_CNT | Developing Count | Numeric  |
| PROFICIENT_CNT | Proficient Count | Numeric  |
| DISTINGUISHED_CNT | Distinguished Count | Numeric  |
| BEGIN_PCT | Beginning Percent | Numeric  |
| DEVELOPING_PCT | Developing Percent | Numeric  |
| PROFICIENT_PCT | Proficient Percent | Numeric  |
| DISTINGUISHED_PCT | Distinguished Percent | Numeric  |

##### Directly Certified

Directly Certified files are retrieved for fiscal years 2017 and 2018, which correspond to 
school calendar years 2017-18 and 2018-19. The file names are based on school calendar years
and the internal dates are fiscal years. Theses file are downloaded manually because
the latest file is not directly downloadable. The files are formatted as Microsoft
Excel files (.xls).

|Variable Name|Variable Label| Data Type |
|-------------|--------------|-----------|
| FISCAL_YEAR | Fiscal Year | Character |
| sys_sch | Concatenation of school district code and institution number | Character |
| SYSTEM_ID | School District Code | Character |
| SYSTEM_NAME | School District Name | Character |
| SCHOOL_ID | Institution Number | Character |
| SCHOOL_NAME | Institution Name | Character |
| direct_cert_perc | Direct certification percent | Numeric |

The sys_sch, SYSTEM_NAME, SCHOOL_NAME fields are ignored.  The SYSTEM_ID
SCHOOL_ID variables are used to construct a institution_id.

### Tidy Data

|Data Category|Path|
|-------------|----|
|Districts | data/districts.csv
|Schools | data/schools.csv |
|School Category (e.g Middle School) | data/schools.csv |
|Grades Served (e.g., KK and 1st) | data/schools.csv |
|Total Enrollment |data/enrollment/totalEnrollment.csv |
|Grade Level Enrollment |data/enrollment/gradeLevel.csv |
|Enrollment by Subgroup |data/enrollment/enrollmentBySubgroup.csv |
|End of Course Milestones |data/milestones/eoc.csv |
|End of Grade Milestones |data/milestones/eog.csv |
|Administrator Salaries and Counts|data/cpi/adminstators.csv|
|Teacher Salaries and Counts|data/cpi/teachers.csv|
|Support Personnel Salaries and Counts|data/cpi/supportPersonnel.csv|
|Years Experience for Administrators, Teachers and Support Personnel | data/cpi/experience.csv|
|Average Years Experience for Administrators, Teachers and Support Personnel | data/cpi/experienceAvg.csv|
|Gender Counts for Administrators, Teachers and Support Personnel | data/cpi/gender.csv|
|Certified Personnel Counts (Provisional or Professional) for Administrators, Teachers and Support Personnel | data/cpi/certifiedPersonnel.csv|
|Certificate Level Counts for Administrators, Teachers and Support Personnel | data/cpi/certificateLevel.csv|
|Race and Ethnicity Counts for Administrators, Teachers and Support Personnel | data/cpi/raceEthnicity.csv|

### Districts

|Variable Name|Variable Label| Data Type |
|-------------|--------------|-----------|
|year|Fiscal Year|integer|
|school_district_id|Three Digit School District ID| character |
|school_district_name| School District Name| character |

### Schools

|Variable Name|Variable Label| Data Type |
|-------------|--------------|-----------|
|year|Fiscal Year|integer|
|school_district_id|Three Digit School District ID| character |
|school_id| School Id (concatination of school_district_id and school_code) | character |
|school_code| School Code (Institution Number)| character |
|school_name| School Name | character |
|PK | Pre-K Served| logical|         
|KK| Kindergarten Served | logical |
|G01| 1st Grade Served| logical |
|G02| 2nd Grade Served| logical |
|G03| 3rd Grade Served| logical |
|G04| 4th Grade Served| logical |
|G05| 5th Grade Served| logical |         
|G06| 6th Grade Served| logical |
|G07| 7th Grade Served| logical |
|G08| 8th Grade Served| logical |
|G09| 9th Grade Served| logical |
|G10| 10th Grade Served| logical |
|G11| 11th Grade Served| logical |
|G12| 12th Grade Served| logical |
|category| School Category (e.g., Middle School) | character |
|period| Fall or Spring| character |
|total|Total Fall Enrollment| integer |

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
