---
title: "Data Explorations"
output: html_notebook
---

# Overview
I have downloaded for types of data:

 * details about certified personnel
 * enrollment numbers both total and by subgroup
 * economic status
 * Georgia Milestone outcomes

For this data exploration, I'm just going to look at Cobb County School District for 2018.
I begin downloading the CPI data using tidyCPI().  This data comes from the 
the Georgia Governor's Office of Student Achievement [Downloadable Data](https://gosa.georgia.gov/report-card-dashboards-data/downloadable-data) site.
The cleaned-up data is found in the "data/cpi" directory.

I'm going to load the CPI data.
```{r}

source("getData.R")
options(dplyr.summarise.inform = FALSE)
library(kableExtra)

tidyCPI()
tidyEnrollment()
tidyEconomic()
```
I'm going to create a table for the institutions in Cobb County.  I'll print
out a summary as a check.
```{r}
districtInstitutions <- getDistrictInstitutions("Cobb County")
summary(districtInstitutions)
```
Looks good.  Next I'll load the the average experience table and construct a
histogram.
```{r}
experienceAvg <- merge(read.csv("data/cpi/experienceAvg.csv"), districtInstitutions) %>%
  filter(year=="2018") %>%
  select(institution_id, employee_type, average)
hist(experienceAvg$average, col="green")
```
I'll do the same thing, but restrict the histogram to teachers.

```{r}
teacherExperienceAvg <- experienceAvg %>%
  filter(employee_type=="PK-12 Teachers") %>%
  select(institution_id, average)
hist(teacherExperienceAvg$average, col="green", main="Average Teacher Experience Per School", 
     xlab="Average Experience")
```
The stark difference (nothing above 18) is the result of administrators; they all have a lot of experience.
The interesting point here is that some institutions have pretty low average experience for teachers.

Next I'll load the the salary table and construct a
histogram. First for administrators.
```{r}
administrators <- merge(read.csv(administratorsDataFile()), districtInstitutions) %>%
  filter(year=="2018")

hist(administrators$avg_annual_salary, col="green", main="Average Administrator Salary Per School", 
     xlab="Average Salary")

```
Next, I'll do the same for teacher salary.
```{r}
teachers <- merge(read.csv(teachersDataFile()), districtInstitutions) %>%
  filter(year=="2018")

hist(teachers$avg_annual_salary, col="green", main="Average Teacher Salary Per School", 
     xlab="Average Salary")

```
There is a low salary school -- where is that.
```{r}
lowSalary <- teachers %>%
  filter(avg_annual_salary < 45000)

lowSalary %>%
  select(institution_name, avg_annual_salary) %>%
  arrange(avg_annual_salary) %>%
  kbl() %>%
  kable_styling()


```
Next, I'll do the same for support personnel salary.
```{r}
support <- merge(read.csv(supportPersonnelDataFile()), districtInstitutions) %>%
  filter(year=="2018")

hist(support$avg_annual_salary, col="green", main="Average Support Personnel Salary Per School", 
     xlab="Average Salary")

```
There are some low salary schools -- where are they.
```{r}
lowSalary <- support %>%
  filter(avg_annual_salary < 50000)

lowSalary %>%
  select(institution_name, avg_annual_salary) %>%
  arrange(avg_annual_salary) %>%
  kbl() %>%
  kable_styling()


```
Next, I'll look at enrollment.  I'll at just the fall.
```{r}
enrollment <- merge(read.csv(enrollmentDataFile()), districtInstitutions) %>%
  filter(year=="2018", period=="Fall") %>%
  group_by(institution_id) %>%
  summarize(total = sum(count))

hist(enrollment$total, col="green", main="Total Enrollment Per Per School", 
     xlab="Enrollment All Grades")
```
Next, I'll look at students per teacher.
```{r}
studentTeacherRatio <- merge(enrollment, teachers) %>%
  mutate(ratio=total/count)

hist(studentTeacherRatio$ratio, col="green", main="Student/Teacher Ratio", 
    xlab="Ratio")

```
There are some schools with high student teacher ratios.
```{r}
highStudentTeacherRatio <- studentTeacherRatio %>%
  filter(ratio > 18)

highStudentTeacherRatio %>%
  select(institution_name, ratio) %>%
  arrange(ratio) %>%
  kbl() %>%
  kable_styling()

```
Next, I'll look at student economic data.  These are direct certified percentages.
```{r}
economic <- merge(read.csv(economicDataFile()), districtInstitutions) %>%
  filter(year=="2018")

hist(economic$direct_certified_percent, col="green", main="Directly Certified Percentage", 
    xlab="Percentage")

```
I'll look at the ones above 40%.
```{r}
highPercent <- economic %>%
  filter(direct_certified_percent > 35)

highPercent %>%
  select(institution_name, percentage=direct_certified_percent) %>%
  arrange(percentage) %>%
  kbl() %>%
  kable_styling()


```



```{r}

lowExperience <- teacherExperienceAvg %>%
  filter(average < 10.0)

teacherSalary <- merge(lowExperience, read.csv(teachersDataFile()) %>%
  filter(year=="2018") %>%
  select(institution_id, avg_annual_salary, count))

administratorSpend <- merge(lowExperience, read.csv(administratorsDataFile()) %>%
  filter(year=="2018") %>%
  select(institution_id, avg_annual_salary, count)) %>%
  group_by(institution_id) %>%
  summarize(totalAdministratorSpend = avg_annual_salary*count)

supportPersonnelSpend <- merge(lowExperience, read.csv(supportPersonnelDataFile()) %>%
  filter(year=="2018") %>%
  select(institution_id, avg_annual_salary, count)) %>%
  group_by(institution_id) %>%
  summarize(totalSupportSpend = avg_annual_salary*count)

teacherSpend <- teacherSalary %>%
  group_by(institution_id) %>%
  summarize(totalTeacherSpend = avg_annual_salary*count)

spend <- merge(administratorSpend, merge(supportPersonnelSpend, teacherSpend)) %>%
  mutate(totalSpend=totalAdministratorSpend+totalSupportSpend+totalTeacherSpend) %>%
  select(institution_id, totalSpend, totalTeacherSpend)

enrollment <- merge(lowExperience, read.csv(enrollmentDataFile()) %>%
  filter(year=="2018", period=="Fall")) %>%
  group_by(institution_id) %>%
  summarize(total = sum(count))

investmentPerStudent <- merge(spend, enrollment) %>%
  mutate(overall_ips = totalSpend/total, teacher_ips=totalTeacherSpend/total) %>%
  select(institution_id, overall_ips, teacher_ips)

economic <- read.csv(economicDataFile()) %>%
  filter(year=="2018")

merge(merge(merge(merge(lowExperience, districtInstitutions), economic), teacherSalary), investmentPerStudent) %>%
  select(institution_name, direct_certified_percent, average, salary=avg_annual_salary, 
         overall_ips, teacher_ips) %>%
  arrange(teacher_ips) %>%
  kbl() %>%
  kable_styling()


```
So the schools with a lot of economic disadvantaged students have pretty low experience levels.  What about the highest level of experience.
```{r}

highExperience <- teacherExperienceAvg %>%
  filter(average > 16.0)

teacherSalary <- merge(highExperience, read.csv(teachersDataFile()) %>%
  filter(year=="2018") %>%
  select(institution_id, avg_annual_salary, count))

administratorSpend <- merge(highExperience, read.csv(administratorsDataFile()) %>%
  filter(year=="2018") %>%
  select(institution_id, avg_annual_salary, count)) %>%
  group_by(institution_id) %>%
  summarize(totalAdministratorSpend = avg_annual_salary*count)

supportPersonnelSpend <- merge(highExperience, read.csv(supportPersonnelDataFile()) %>%
  filter(year=="2018") %>%
  select(institution_id, avg_annual_salary, count)) %>%
  group_by(institution_id) %>%
  summarize(totalSupportSpend = avg_annual_salary*count)

teacherSpend <- teacherSalary %>%
  group_by(institution_id) %>%
  summarize(totalTeacherSpend = avg_annual_salary*count)

spend <- merge(administratorSpend, merge(supportPersonnelSpend, teacherSpend)) %>%
  mutate(totalSpend=totalAdministratorSpend+totalSupportSpend+totalTeacherSpend) %>%
  select(institution_id, totalSpend, totalTeacherSpend)

enrollment <- merge(highExperience, read.csv(enrollmentDataFile()) %>%
  filter(year=="2018", period=="Fall")) %>%
  group_by(institution_id) %>%
  summarize(total = sum(count))

investmentPerStudent <- merge(spend, enrollment) %>%
  mutate(overall_ips = totalSpend/total, teacher_ips=totalTeacherSpend/total) %>%
  select(institution_id, overall_ips, teacher_ips)

economic <- read.csv(economicDataFile()) %>%
  filter(year=="2018")

merge(merge(merge(merge(highExperience, districtInstitutions), economic), teacherSalary), investmentPerStudent) %>%
  select(institution_name, direct_certified_percent, average, avg_annual_salary, 
         overall_ips, teacher_ips) %>%
  arrange(teacher_ips) %>%
  kbl() %>%
  kable_styling()

```

So the schools with the highest levels of experience have twice that of those with the lowest levels of experience.
Moreover, their students tend not to be economically disadvanaged.
```{r}
midExperience <- teacherExperienceAvg %>%
  filter(average > 10.0, average < 15)

teacherSalary <- merge(midExperience, read.csv(teachersDataFile()) %>%
  filter(year=="2018") %>%
  select(institution_id, avg_annual_salary, count))

administratorSpend <- merge(midExperience, read.csv(administratorsDataFile()) %>%
  filter(year=="2018") %>%
  select(institution_id, avg_annual_salary, count)) %>%
  group_by(institution_id) %>%
  summarize(totalAdministratorSpend = avg_annual_salary*count)

supportPersonnelSpend <- merge(midExperience, read.csv(supportPersonnelDataFile()) %>%
  filter(year=="2018") %>%
  select(institution_id, avg_annual_salary, count)) %>%
  group_by(institution_id) %>%
  summarize(totalSupportSpend = avg_annual_salary*count)

teacherSpend <- teacherSalary %>%
  group_by(institution_id) %>%
  summarize(totalTeacherSpend = avg_annual_salary*count)

spend <- merge(administratorSpend, merge(supportPersonnelSpend, teacherSpend)) %>%
  mutate(totalSpend=totalAdministratorSpend+totalSupportSpend+totalTeacherSpend) %>%
  select(institution_id, totalSpend, totalTeacherSpend)

enrollment <- merge(midExperience, read.csv(enrollmentDataFile()) %>%
  filter(year=="2018", period=="Fall")) %>%
  group_by(institution_id) %>%
  summarize(total = sum(count))

investmentPerStudent <- merge(spend, enrollment) %>%
  mutate(overall_ips = totalSpend/total, teacher_ips=totalTeacherSpend/total) %>%
  select(institution_id, overall_ips, teacher_ips)

economic <- read.csv(economicDataFile()) %>%
  filter(year=="2018")

merge(merge(merge(merge(midExperience, districtInstitutions), economic), teacherSalary), investmentPerStudent) %>%
  select(institution_name, direct_certified_percent, average, salary=avg_annual_salary, 
         overall_ips, teacher_ips) %>%
  arrange(teacher_ips) %>%
  kbl() %>%
  kable_styling()

```

