---
title: "Data Explorations"
output: html_notebook
---

# Overview
I have downloaded for types of data:

 * details about certified personnel
 * enrollment numbers both total and by subgroup
 * economic status
 * Georgia Milestone outcomes

For this data exploration, I'm just going to look at Cobb County School District for 2018.
I download and cleanup CPI, enrollment, and milestone data.  I was using direcly
certified data, but I have decided to use economically disadvantaged.
All data comes from the 
the Georgia Governor's Office of Student Achievement [Downloadable Data](https://gosa.georgia.gov/report-card-dashboards-data/downloadable-data) site.
The cleaned-up data is found in the "data" directory in cpi, enrollment, and
milestone folders.

As a check, I'm going to create a table for the institutions in Cobb County.  I'll print
out a summary as a check.  As a side effect, I will have a list of Cobb Schools.
I've exclude Devereux Ackerman Academy and Cobb Horizon High School because
these are unique types of schools.  It may be appropriate to exclude Kennesaw
Charter as well.

```{r}

source("getData.R")
source("features.R")
options(dplyr.summarise.inform = FALSE)
library(kableExtra)

tidyCPI()
tidyEnrollment()
tidyEnrollmentBySubgroup()
tidyMilestoneEoC()
tidyMilestoneEoG()

cobbSchools <- getDistrictInstitutions("Cobb County", "2018") %>%
  filter(institution_name!="Devereux Ackerman Academy", institution_name!="Cobb Horizon High School",
         institution_name != "Kennesaw Charter School")

getDistrictSummary(cobbSchools) %>%
  select("Category"=it, "Count"=it_count) %>%
  kbl() %>%
  kable_styling()

```
Looks good.  Next I'll load the the average experience table and construct a
histogram.
```{r}
avgExp <- getAverageExperience("2018", institutions=cobbSchools) %>%
  select(institution_id, employee_type, elementary, middle, high, average)
par(mfrow=c(2,2))

plotAvgExp <- function(data, title) {
  with(data, hist(average, xlab="Average Experience", main=title))
}

plotAvgExp(subset(avgExp, employee_type=="Administrators"), "Administrators")
plotAvgExp(subset(avgExp, employee_type=="PK-12 Teachers" & elementary), "Elementary School Teachers")
plotAvgExp(subset(avgExp, employee_type=="PK-12 Teachers" & middle), "Middle School Teachers")
plotAvgExp(subset(avgExp, employee_type=="PK-12 Teachers" & high), "High School Teachers")
```
The stark difference (nothing above 18) is the result of administrators; they all have a lot of experience.
The interesting point here is that some institutions have pretty low average experience for teachers.

Next I'll do the same thing for average annual salary.
```{r}
adminPay <- getAdministrators(2018, cobbSchools)
teacherPay <- getTeachers(2018, cobbSchools)

par(mfrow=c(2,2))

plotAvgSalary <- function(data, title) {
  with(data, hist(avg_annual_salary, xlab="Average Annual Salary", main=title, breaks=8))
}

plotAvgSalary(adminPay, "Administrators")
plotAvgSalary(subset(teacherPay, elementary), "Elementary School Teachers")
plotAvgSalary(subset(teacherPay, middle), "Middle School Teachers")
plotAvgSalary(subset(teacherPay, high), "High School Teachers")

```
There is a low salary school -- where is that.
```{r}
lowSalary <- teachers %>%
  filter(avg_annual_salary < 45000)

lowSalary %>%
  select(institution_name, avg_annual_salary) %>%
  arrange(avg_annual_salary) %>%
  kbl() %>%
  kable_styling()
```
Next, I'll look at fall enrollment.
```{r}
enrollment <- getEnrollment(2018, cobbSchools)
esEnrollment <- merge(subset(enrollment, elementary), cobbSchools)
msEnrollment <- merge(subset(enrollment, middle), cobbSchools)
hsEnrollment <- merge(subset(enrollment, high), cobbSchools)

par(mfrow=c(2,2))

plotEnrollment <- function(data, title) {
  with(data %>%
         group_by(institution_id) %>%
         summarize(total = sum(count)),
       hist(total, xlab="Total Enrollment", main=title, breaks=6))
}

plotEnrollment(esEnrollment, "Elementary Schools")
plotEnrollment(msEnrollment, "Middle Schools")
plotEnrollment(hsEnrollment, "High Schools")
```
Next, I'll look at students per teacher.
```{r}
par(mfrow=c(2,2))

plotStudentTeacherRatio <- function(enrollment, teachers, title) {
  with(merge(enrollment %>%
               group_by(institution_id) %>%
               summarize(total = sum(count)),
             teachers) %>%
         mutate(ratio=total/count),
       hist(ratio, xlab="Student/Teacher Ratio", main=title,
            breaks=c(8, 10, 12, 14, 16, 18, 20, 22)))
}

plotStudentTeacherRatio(esEnrollment, teachers, "Elementary Schools")
plotStudentTeacherRatio(msEnrollment, teachers, "Middle Schools")
plotStudentTeacherRatio(hsEnrollment, teachers, "High Schools")
```
There are some schools with low student teacher ratios.
```{r}
par(mfrow=c(1,3))

enrollmentBySubgroup <- getEnrollmentBySubgroup(2018, cobbSchools)

plotEnrollmentBySubgroup <- function(enrollment, title) {
  with(enrollment, {
    count <- nrow(enrollment)
    title <- sprintf("%s (%d)", title, count)
    hist(enroll_percent_ed, xlab="Percent Economically Disadvantaged",
         main=title, breaks=c(0, 25, 50, 75, 100))
  })
}

plotEnrollmentBySubgroup(subset(enrollmentBySubgroup, elementary), "Elementary Schools")
plotEnrollmentBySubgroup(subset(enrollmentBySubgroup, middle), "Middle Schools")
plotEnrollmentBySubgroup(subset(enrollmentBySubgroup, high), "High Schools")
```
Next we will look at milestones for English language arts for 3rd, 4th, and 5th grades.
```{r}
eogEla345Milestones <- getMilestoneEoG(2018, cobbSchools) %>%
  filter(acdmc_lvl %in% c("3", "4", "5"), 
         subgroup_name=="All Students", 
         test_cmpnt_typ_nm=="English Language Arts") %>%
  mutate(ela_score = developing_pct + proficient_pct + distinguished_pct) %>%
  mutate(acdmc_lvl=as.factor(acdmc_lvl))

par(mfrow=c(2,2))

plotEogElaMilestones <- function(milestones, lvl, title) {
  with(milestones %>% filter(acdmc_lvl==lvl),
       hist(ela_score, xlab="ELA Developing + Proficient + Distinguished", ylim=c(0, 30), 
            main=title, breaks = c(50, 60, 70, 80, 90, 100)))
}

plotEogElaMilestones(eogEla345Milestones, "3", "3rd Grade")
plotEogElaMilestones(eogEla345Milestones, "4", "4th Grade")
plotEogElaMilestones(eogEla345Milestones, "5", "5th Grade")
```


```{r}

eogEla345MilestonesWithSubgroups <- merge(eogEla345Milestones, enrollmentBySubgroup, by=c("institution_id"))

qplot(enroll_percent_lep, ela_score, data=eogEla345MilestonesWithSubgroups, 
      facets=.~acdmc_lvl, geom=c("point", "smooth"), 
      ylab="ELA Developing + Proficient + Distinguished",
      xlab="Percent Limited English Proficiency")

```

```{r}
qplot(acdmc_lvl, ela_score, data=eogEla345MilestonesWithSubgroups, 
      geom="boxplot", 
      ylab="ELA Developing + Proficient + Distinguished",
      xlab="Grade")

```

LEP versus ELA


Years of experience versus econimically disadvantaged.

```{r}

teacherAvgExp <- subset(avgExp, employee_type=="PK-12 Teachers" & elementary)

teacherAvgExp_enrollmentBySubgroup <- merge(enrollmentBySubgroup, teacherAvgExp)

qplot(enroll_percent_lep, average, data=teacherAvgExp_enrollmentBySubgroup, 
      ylab="Average Teacher Experience",
      xlab="Percent Limited English Proficiency") + geom_smooth(method = "lm")

```
```{r}

eogEla345MilestonesWithAvgExp <- merge(eogEla345Milestones, teacherAvgExp, by=c("institution_id"))


qplot(average, ela_score, data=eogEla345MilestonesWithAvgExp, 
      facets=.~acdmc_lvl, 
      ylab="ELA Developing + Proficient + Distinguished",
      xlab="Average Teacher Experience") + geom_smooth(method = "lm")

```

```{r}

studentTeacherRatio <- merge(enrollment %>%
                               group_by(institution_id) %>%
                               summarize(total = sum(count)),
                             teachers) %>%
  mutate(ratio=total/count)
eogEla345MilestonesWithStudentTeacherRatio <- merge(eogEla345Milestones, studentTeacherRatio, by=c("institution_id"))

qplot(ratio, ela_score, data=eogEla345MilestonesWithStudentTeacherRatio, 
      facets=.~acdmc_lvl, 
      ylab="ELA Developing + Proficient + Distinguished",
      xlab="Student/Teacher Ratio") + geom_smooth(method = "lm")


```

```{r}
studentTeacherRatio <- merge(enrollment %>%
                               group_by(institution_id) %>%
                               summarize(total = sum(count)),
                             teachers) %>%
  filter(elementary) %>%
  mutate(ratio=total/count)
enrollmentBySubgroupWithStudentTeacherRatio <- merge(enrollmentBySubgroup, studentTeacherRatio, by=c("institution_id"))

qplot(ratio, enroll_percent_ed, data=enrollmentBySubgroupWithStudentTeacherRatio, 
      ylab="Percent Economically Disadvantaged",
      xlab="Student/Teacher Ratio") + geom_smooth(method = "lm")

```

What I am seeing is that schools can control student/teacher ratio and it is 
generally related to the percent of student population that is economically
disadvantaged, but there is a lot of variation.  I would say too much.  On the
other hand, teacher experience seems to be positively correlated with test
outcomes.  I suspect schools have little control over experience.  The choice
to move is a teacher's choice.  Nonetheless, I might be able to demonstrate that
test_score ~ teacher_experience + student_teacher_ratio + pct_lep.

